2.4 PRNT50

### 2024-05-27 2_scripts/2.4 PRNT50.qmd
Anti-S sensitivity analysis update

p09e
```

>>>>

p09e
```
## 4 Misc Analysis

#### 4.1 Sensitivity Analysis
```{r}
# Add vax variable
APLs<-merge(APLs, select(RFD4682_e, order_ID, VACCINATION_STATUS),
      by="order_ID", all=T)

APLs$vax<-APLs$VACCINATION_STATUS
APLs<-select(APLs, -c("VACCINATION_STATUS"))
table(APLs$vax, useNA = 'ifany')

# Add days since imm (vax)
APLs<-merge(APLs, select(RFD4682_e, order_ID, DAYS_SINCE_IMM),
      by="order_ID", all=T)
```

```{r}
# re-do the melt data
tb_abbsv<-APLs %>%
  .[.$abb_s_AU==1 & !(.$vax %in% c("Not Immunized", "Pre-Immunization")) &
    .$DAYS_SINCE_IMM>=28 & !is.na(.$DAYS_SINCE_IMM),] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott_s= sum(pos_abbott_s, na.rm = T), 
            pos_abbott_s2= sum(pos_abbott_s2, na.rm = T)
            )
tb_abbsv$prop.abbs1<-tb_abbsv$pos_abbott_s/tb_abbsv$n
tb_abbsv$prop.abbs2<-tb_abbsv$pos_abbott_s2/tb_abbsv$n

# clean NAs
tb_abbsv<-tb_abbsv[!is.na(tb_abbsv$samplemonth),]

tb_abbsv_melt<-reshape2::melt(select(tb_abbsv, -c(2:4)), 
                        id.vars = c("samplemonth"),
                        variable.name = "COI",
                        value.name = "prop")
```

```{r}
# re-do the plot
# APL vax only
p09v0<-tb_abbsv_melt %>%
  ggplot(aes(x=samplemonth, y=prop, group=COI, colour=COI)) +
  geom_point(aes(colour=COI)) +
  geom_line(aes(colour=COI))  +
  theme_bw() + 
  ylim(c(0,1))
p09v0
```
## 4.2 The vaccination rate among APL
The APL data Carmen shared with us has vaccination info, the 2024 Jan ver of it even has the vax brand. While the CBS data we have doesnot include vax info. However, a paper written by Sheila has the vax% of those in CBS data, which we can use for comparison on vax%.

```{r}
# create vax% cat var
APLs$vax_cat<-rep(NA, 214780)
APLs$vax_cat[APLs$vax %in% c("1 Dose","1-2 Doses","2 Doses","3 Doses",
                             "4 Doses")]<-1
APLs$vax_cat[APLs$vax %in% c("Not Immunized", "Pre-Immunization")]<-0
table(APLs$vax, APLs$vax_cat, useNA = 'ifany')
```
```{r}
# create merged data
tb_vax<-APLs %>%
  .[.$abb_s_AU==1,] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            n_vax= sum(vax_cat, na.rm = T)
            )
tb_vax$prop.vax<-tb_vax$n_vax/tb_vax$n

# clean NAs
tb_vax<-tb_vax[!is.na(tb_vax$samplemonth),]

# make the plot
p10<-tb_vax %>%
  ggplot(aes(x=samplemonth, y=prop.vax)) +
  geom_point() +
  geom_line()  +
  geom_smooth(colour="red") +
  theme_bw() + 
  ylim(c(0,1)) +
  ylab("Vaccinated%")
p10
```
**Alton's comments:** "It might be worth comparing to reported vax rate for Alberta similar to what Sheila and Steve did in that figure." As they used (S+ N-)% to estimate the vax%, it will inevitably results in underestimation, as those who are vaccinated and infected was not counted.

```{r}
# create an addition indicator for S+ and N-
table(APLs$pos_abbott, APLs$pos_abbott_s, useNA = 'ifany')

APLs$vax_est<-rep(0, 214780)
APLs$vax_est[APLs$pos_abbott==0 & APLs$pos_abbott_s==1]<-1
APLs$vax_est[is.na(APLs$pos_abbott) | is.na(APLs$pos_abbott_s)]<-NA
table(APLs$vax_est)
```
```{r}
tb_vax_est<-APLs %>%
  .[.$abb_s_AU==1 & !is.na(.$pos_abbott_s) & !is.na(.$pos_abbott),] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            n_vax= sum(vax_est, na.rm = T)
            )
tb_vax_est$prop.vax<-tb_vax_est$n_vax/tb_vax_est$n

# clean NAs
tb_vax_est<-tb_vax_est[!is.na(tb_vax_est$samplemonth),]

# make the plot
p11<-tb_vax_est %>%
  ggplot(aes(x=samplemonth, y=prop.vax)) +
  geom_point() +
  geom_line()  +
  geom_smooth(colour="red") +
  theme_bw() + 
  ylim(c(0,1)) +
  ylab("Vaccinated% Est by anti-S&N")
p11
```
